---
title: "Tutorial: Functional Data Analysis (FDA)"
author: "matilde trevisani"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    fig_width: 8
    fig_height: 5
    df_print: "tibble" # other options: paged, kable, default
  # bookdown::html_document2:
  #   fig_caption: yes
  #   number_sections: false
  #   base_format: rmarkdown::html_vignette
  # bookdown::pdf_document2:
  #   fig_caption: yes
  #   number_sections: false
   # highlight: espresso    # textmate  breezedark  haddock  espress monochrome  pygments
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Tutorial: Functional Data Analysis (FDA)}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r optcmd, echo=FALSE, cache=F}
draft <- T
track <- T
figeval <- T
raw <- T
over <- T
```

```{r utils, echo=F, cache=F}
nam_work <- "JPSP"
namtd <- "jpsp"
normty <- "nchi"
normtytxt <- "chi square-like"
```

```{r setup, include=FALSE, cache=F}
knitr::opts_chunk$set(
  echo = track, 
  collapse = TRUE, 
  comment = "#>", 
  fig.path = paste0(nam_work,'_fig/fda/',normty,'_'), 
  fig.lp=paste0(normty,'_'),
  cache = 2, 
  cache.path=paste0(namtd,'_fda_',normty,'_cache/'), 
  cache.comments = F, 
  autodep = TRUE, 
  cache.extra = knitr::rand_seed,
  par = TRUE, # ??
  fig.align = 'center',
  crop = TRUE, # ??
  dev = c('CairoPNG','CairoPDF'), # ??
  fig.width = 8,
  fig.height = 5,
  out.width = '100%',
  warning = track,
  message = track
  )

# Cairo opzionale per figure di alta qualità
if (requireNamespace("Cairo", quietly = TRUE)) {
  knitr::opts_chunk$set(list(png = list(type = "cairo"), pdf = list(type = "cairo")))
}

knitr::opts_knit$set(aliases=c(h = 'fig.height', w = 'fig.width'), unnamed.chunk.label = "fig")


knitr::knit_hooks$set(par=function(before, options, envir){
  if (before && options$fig.show!='none') {
    par(mar=c(2.,1.9,.25,.25)*sqrt(options$fig.width/6),mgp=c(1.1,.3,0)*sqrt(options$fig.width/6),tcl=-.15*sqrt(options$fig.width/6),font.main=1,cex.lab=.75*sqrt(options$fig.width/6),cex.axis=.8*sqrt(options$fig.width/6),cex.main=sqrt(options$fig.width/6),xaxs="i",yaxs="i")#default family is sans
  }  
})
```

```{r packages, include=FALSE, cache=F}
# Main package
library(cccc)

# Data manipulation
library(dplyr)
library(tibble) 

# Visualization
library(ggplot2)
#library(patchwork)
library(ggpubr)

# Interactive tables
library(DT)

library(fda)
library(lattice)  # for levelplot
library(TeachingDemos)  # for subplot()
```

# FDA

## Normalization type: `r normtytxt`

```{r internaldat, echo=F, results='hide', eval=raw}
# loading the dataset is necessary only when knitting from the Rstudio.
# instead, datnetcetc is found (as internal dataset) when building vignettes from within the package.
#load(paste0("../data-raw/dat",normty,".rda"))
sysdata_filenames <- load("../R/sysdata.rda")
sysdata_filenames
```

Data consists of the time by term table of normalized frequencies (since
this is the format required as input to `fda` routines).

```{r fdautils, echo=track, include=draft, eval=T}
if(exists(paste0("dat", normty))) {
  cat("Dataset 'dat", normty, "' found in internal data.\n", sep = "")
} else {
  cat("Dataset 'dat", normty, "' not found in internal data.\n We create it.", sep = "")
  tdm <- system.file("extdata", "tdm.csv", package = "cccc")
  corpus <- system.file("extdata", "corpus.csv", package = "cccc")
  data <- importData(tdm_file = tdm, corpus_file = corpus)
  datnchi <- normalization(data, normty = "nchi", sc = 1)
}

dat <- get(paste0("dat", normty))
ind_d <- dat$year_cols
names(dat)
n_zones <- length(dat$zone)
dat <- dat$tdm
dat %>% dplyr::select(1:ind_d[1]-1,ind_d[1:5])%>% head 
year <- names(dat)[ind_d]
keyword <- as.character(dat$keyword)
mat <- t(dat[,ind_d])
dimnames(mat) <- list(year,keyword)
cat(nt <- nrow(mat), "time points\n")
cat(n <- ncol(mat), "terms\n")
cat("A subset:")
mat[c(1:3,(nt-2):nt),c(1:3,(n-2):n)]
```

A *cubic spline smoothing*, fitted by an ordinary least squares
estimate, is taken as the starting point. Note that we merely obtain an
exact interpolation of data, without any smoothing effect.

```{r curvespar, include=track, eval=T}
ylab <- paste0("word (",normtytxt,") normalized frequency")
r <- 2
samplpo <- year %>% as.numeric
ny <- diff(range(samplpo))+1 
xaxlab <- (samplpo)[1]+0:(ny-1)
xaxt <- seq(1, ny, by=r)
xaxlab[-xaxt] <- ""

xnug <- 100; xlim <- c(1,ny)+c(-1,1)*ny/xnug
ynug <- 20; ylim <- range(mat)+c(-1,1)*diff(range(mat))/ynug

lwd.u <- .35; lwd.bty.u <- 1
col.tck.u <- "slategray4"; col.bty.u <- "slategray3"
```

The figure below show the fit of a cubic smoothing spline to the `r normtytxt` normalized data for a few selected words (top) and for all
words (bottom).
<!-- \@ref(fig:sat)  -->

```{r sat, w=8, h=4.65, fig.cap=paste("Fit of a cubic smoothing spline to", normtytxt," normalized data."), fig.pos="p", fig.show='hold', echo=track, results='hide'}

fdtime <- 1+samplpo-samplpo[1]
br <- fdtime 
nord <- 4
bbasis <- create.bspline.basis(norder=nord, breaks=br)

# For selected words
k <- match(kw_list[seq(1, le=n_zones, by=3)], keyword)
# dimnames(mat)[[2]][k]
# keyword[k]

sat <-Data2fd(fdtime, mat[, k], bbasis, fdnames = list("time", "words", "observations"))

curves <- sat

plot(curves, xlim=xlim, ylim=ylim, nx=ny, xaxt="n", xlab="", ylab=ylab, lwd=lwd.u*2, axes=F) 
mtext(paste("words:",paste(keyword[k],collapse=", ")), 3, line=-1)
axis(1, at=1:ny, labels = xaxlab, las=2, col.ticks = col.tck.u)
axis(2)
box(col=col.bty.u, lwd=lwd.bty.u)

matpoints(1+samplpo-samplpo[1], mat[,c(k)], pch=20, cex=.75)

# For all words
sat <-Data2fd(fdtime, mat, bbasis, fdnames = list("time", "words", "observations"))

curves <- sat

plot(curves, xlim=xlim, ylim=ylim, nx=nt, xaxt="n", xlab="", ylab=ylab, lwd=lwd.u, axes=F)
axis(1, at=1:ny, labels = xaxlab, las=2, col.ticks = col.tck.u)
axis(2)
box(col=col.bty.u, lwd=lwd.bty.u)
```

```{r my_plotfd, include=F}
my_plotfd <- function (x, y, Lfdobj = 0, href = TRUE, titles = NULL, xlim = NULL,
                    ylim = NULL, xlab = NULL, ylab = NULL, ask = FALSE, nx = NULL,
                    axes = NULL, ...)
{
  fdobj <- x
  if (!(is.fd(fdobj) || is.fdPar(fdobj)))
    stop(paste("First argument is neither a functional data or a ",
               "functional parameter object."))
  if (is.fdPar(fdobj))
    fdobj <- fdobj$fd
  if (is.null(axes)) {
    if (is.null(fdobj$basis$axes)) {
      Axes <- TRUE
      axFun <- FALSE
    }
    else {
      if (!inherits(fdobj$basis$axes, "list"))
        stop("fdobj$basis$axes must be a list;  ", "class(fdobj$basis$axes) = ",
             class(fdobj$basis$axes))
      if (!(inherits(fdobj$basis$axes[[1]], "character") ||
            inherits(fdobj$basis$axes[[1]], "function")))
        stop("fdobj$basis$axes[[1]] must be either a function or the ",
             "name of a function;  class(fdobj$basis$axes[[1]]) = ",
             class(fdobj$basis$axes[[1]]))
      Axes <- FALSE
      axFun <- TRUE
      axList <- c(fdobj$basis$axes, ...)
    }
  }
  else {
    if (is.logical(axes)) {
      Axes <- axes
      axFun <- FALSE
    }
    else {
      if (!inherits(axes, "list"))
        stop("axes must be a logical or a list;  class(axes) = ",
             class(axes))
      if (!(inherits(axes[[1]], "character") || inherits(axes[[1]],
                                                         "function")))
        stop("axes[[1]] must be either a function or the ",
             "name of a function;  class(axes[[1]]) = ",
             class(axes[[1]]))
      Axes <- FALSE
      axFun <- TRUE
      axList <- c(axes, ...)
    }
  }
  Lfdobj <- int2Lfd(Lfdobj)
  if (!inherits(Lfdobj, "Lfd"))
    stop("Second argument is not a linear differential operator.")
  coef <- fdobj$coefs
  coefd <- dim(coef)
  ndim <- length(coefd)
  nbasis <- coefd[1]
  if (is.null(nx))
    nx <- max(c(501, 10 * nbasis + 1))
  nrep <- coefd[2]
  if (ndim > 2)
    nvar <- coefd[3]
  else nvar <- 1
  basisobj <- fdobj$basis
  rangex <- basisobj$rangeval
  if (missing(y)) {
    y <- nx
  }
  else {
    if (is.numeric(y))
      y <- as.vector(y)
  }
  Y <- y
  if (length(y) == 1) {
    if (y >= 1) {
      y <- seq(rangex[1], rangex[2], len = round(y))
    }
    else {
      stop("'y' a single number less than one.")
    }
  }
  if (min(y) < rangex[1] || max(y) > rangex[2])
    stop("Values in Y are outside the basis range.")
  if (is.null(xlim)) {
    xlim <- rangex
  }
  else {
    rangex[1] <- max(rangex[1], xlim[1])
    rangex[2] <- min(rangex[2], xlim[2])
    if (length(Y) == 1)
      y <- seq(rangex[1], rangex[2], len = round(Y))
  }
  fdmat <- eval.fd(y, fdobj, Lfdobj)
  rangey <- range(fdmat)
  if (is.null(ylim))
    ylim <- rangey
  fdnames = fdobj$fdnames
  fdlabelslist = fdlabels(fdnames, nrep, nvar)
  xlabel = fdlabelslist$xlabel
  ylabel = fdlabelslist$ylabel
  casenames = fdlabelslist$casenames
  varnames = fdlabelslist$varnames
  if (is.null(xlab))
    xlab <- xlabel
  if (is.null(ylab))
    ylab <- ylabel
  if (ndim < 2) {
    plot(y, fdmat, type = "l", xlim = xlim, ylim = ylim,
         xlab = xlab, ylab = ylab, axes = Axes, ...)
    if (axFun)
      do.call(axList[[1]], axList[-1])
    if (zerofind(fdmat) && href)
      abline(h = 0, lty = 2)
  }
  if (ndim == 2) {
    if (!ask) {
      matplot(y, fdmat, type = "l", xlim = xlim, ylim = ylim, xlab = xlab, ylab = ylab, axes = Axes, ...)
      if (axFun)
        do.call(axList[[1]], axList[-1])
      if (zerofind(fdmat) && href)
        abline(h = 0, lty = 2)
    }
    else {
      op <- par(ask = FALSE)
      on.exit(par(op))
      cat("Multiple plots:  Click in the plot to advance to the next")
      for (irep in 1:nrep) {
        plot(y, fdmat[, irep], type = "l", xlim = xlim,
             ylim = ylim, xlab = xlab, ylab = ylab, axes = Axes,
             ...)
        if (axFun)
          do.call(axList[[1]], axList[-1])
        if (irep < 2)
          par(ask = ask)
        if (!is.null(casenames))
          title(casenames[irep])
        else title(paste("Case", irep))
        if (zerofind(ylim) && href)
          abline(h = 0, lty = 2)
      }
    }
  }
  if (ndim == 3) {
    if (!ask) {
      for (ivar in 1:nvar) {
        matplot(y, fdmat[, , ivar], type = "l", xlim = xlim, ylim = ylim, xlab = xlab, ylab = ylab, ask = FALSE, axes = Axes, ...)
        if (axFun)
          do.call(axList[[1]], axList[-1])
        if (!is.null(varnames))
          title(varnames[ivar])
        else title(paste("Variable", ivar))
        if (zerofind(ylim) && href)
          abline(h = 0, lty = 2)
      }
    }
    else {
      op <- par(ask = FALSE)
      on.exit(par(op))
      cat("Multiple plots:  Click in the plot to advance to the next")
      for (irep in 1:nrep) {
        for (ivar in 1:nvar) {
          plot(y, fdmat[, irep, ivar], type = "l", xlim = xlim,
               ylim = ylim, xlab = xlab, ylab = ylab, axes = Axes,
               ...)
          if (axFun)
            do.call(axList[[1]], axList[-1])
          if (!is.null(casenames))
            titlestr = casenames[irep]
          else titlestr = paste("Case", irep)
          if (!is.null(varnames)) {
            titlestr = paste(titlestr, "  ", varnames[ivar])
          }
          else {
            titlestr = paste(titlestr, "  ", "Variable",
                             ivar)
          }
          title(titlestr)
          if (zerofind(ylim) && href)
            abline(h = 0, lty = 2)
        }
      }
    }
  }
  "done"
}

```

We can visualize the mean and the standard deviation functions.
<!-- (Figure \ref{nchi_meanEtsd}). --> <!-- \ref{nM_meanEtsd}). -->

```{r meanEtsd, dependson="curvespar", w=8, h=2.75, fig.cap=paste("Mean and Standard Deviation functions (",normtytxt," normalized data).",sep=""), fig.pos="p", fig.show='hold', echo=track, results="hide", eval=T}

### summarizing functional data

# the mean function
mfd <- mean.fd(curves)

my_plotfd(curves, nx=ny, xlim=xlim, ylim=ylim, col="powderblue", xaxt="n", xlab="", ylab=ylab)
axis(1, at=1:ny, labels = xaxlab, las=2)

lines(mfd, nx=ny)

def.par <- par(no.readonly = TRUE) 
nf=layout(matrix(c(rep(0,4),c(0,1,1,0),rep(0,4),rep(0,4)),byrow=T, ncol=4),heights = c(1,6,3,3),widths = c(.25,3,3,.25))
par(new = TRUE, oma = par("mar"), fg=gray(.6), col.axis=gray(.55))

plot(mfd, nx=ny, xlim=xlim, ylim=ylim,
   xaxt="n",xlab="",ylab="",col="Darkslateblue",lwd=2)
axis(1,at=1:ny, labels = xaxlab, las=2)

par(def.par)

# We can also look at the variation in the function as a pointwise process
stdfd <- std.fd(curves)

plot(stdfd, xlim=xlim, xaxt="n", xlab="", ylab="Standard Deviation",yaxs="r", axes=F)
axis(1, at=1:ny, labels = xaxlab, las=2)
axis(2)
```

Covariance and correlation functions may suggest any bundles of curves
temporally related in specific time windows.
<!-- %>%(Figure \ref{nchi_corr}). -->

```{r corr, w=9, h=4, dependson="sat", fig.cap=paste("Covariance and correlation functions (",normtytxt," normalized data).",sep=""), fig.pos="p", fig.show='hold', warning=F, echo=track}

varfd <- var.fd(curves)
varvals <- eval.bifd(1:ny, 1:ny, varfd)

# We can also look at the correlation

corfd <- cor.fd(1:ny,curves)

def.par <- par(mfrow=c(1,2), cex.axis=.7, cex.lab=.85, tcl=-.25)

xylab <- "year" 
  
persp(1:ny, 1:ny, varvals, theta=40, phi=20, expand = .5, ticktype='detailed', nticks =14, col="orange1", xlab=xylab, ylab=xylab, zlab="covariance") # r=3, shade=.2,

persp(1:ny, 1:ny, corfd, theta=40, phi=20, expand = .5, ticktype='detailed', nticks =14, col="yellow1",xlab=xylab, ylab=xylab, zlab="correlation")

par(def.par)

#contour(fdtime, fdtime, varvals)

colRP=c("chartreuse4","white","deepskyblue4")
mv <- max(abs(varvals))
mc <- max(abs(corfd))
rl <- 4; rt <- 2

atxr <- seq(1,ny,by=rt)
labelsxr <- (as.numeric(samplpo[1])+0:(ny-1))[atxr]
xaxt <- seq(1, ny, by=rl)
labelsxr[-xaxt] <- ""

cexax <- .85; tck <- .75; cexlab <- .95

m <- mv
p1 <- levelplot(x=varvals,contour=TRUE,labels=FALSE,scales=list(at=atxr,labels=labelsxr,x=list(rot=90),cex=cexax,tck=tck),at=do.breaks(c(-m,m),21),col.regions=colorRampPalette(colRP),xlab=list(xylab,cex=cexlab),ylab=list(xylab,cex=cexlab),colorkey=list(width=.75))

m <- mc
p2 <- levelplot(x=corfd,scales=list(labels=labelsxr,at=atxr,x=list(rot=90),cex=cexax,tck=tck),at=do.breaks(c(-m,m),21),col.regions=colorRampPalette(colRP),xlab=list(xylab,cex=cexlab),ylab=list(xylab,cex=cexlab),colorkey=list(width=.75))

wi <- 2.75
print(p1, position = c(0, 0, 0.5, 1), more = TRUE, panel.width=list(x=wi, units="in"), panel.height=list(x=wi-1.,units="in"))
print(p2, position = c(0.5, 0, 1, 1), panel.width=list(x=wi, units="in"), panel.height=list(x=wi-1.,units="in"))

```

## Smoothing selection

By the roughness penalty approach for estimation, we smoothed the data
by trying different spline orders ($m$ from $1$ to $8$), with
$\text{PEN}_{m-2}(x) =\int[D^{m-2}x(s)]^2 ds$ as a roughness penalty,
and varying the smoothing parameter over an opportune range of values
($\log_{10}\lambda$ from $-6$ to $9$).
<!-- (Figure \ref{nchi_selection}). --> <!-- \ref{nchi_selection}). -->

If the aim is comparing different normalizations, we can create an object apt to contain the selection output for the all types of normalizations under examination.
(However, if you are solely interested in the selection output for the current normalization, you can skip this preliminary setting.)


### PEN$_{m-2}$


```{r selection_m2, fig.show='hold', fig.cap=paste("Smoothing selection using PEN$_{m-2}$ (",normtytxt," normalized data).",sep=""), fig.pos="p", w=4, h=3, echo=track, warning=F}

p_m2 <- smoothingSelection(
                      datnchi,
                      lambda_seq = NULL,
                      degrees = NULL,
                      penalty_type = "m-2",
                      plot = TRUE,
                      verbose = TRUE)
```

```{r optpar_m2, dependson="selection_m2", echo=track}
names(p_m2)
# options(tibble.print_min = Inf)
# p_m2$results
p_m2$summary_optimal
p_m2$optimal_gcv
p_m2$optimal_ocv
p_m2$opt_order
p_m2$penalty_type

p_m2$plots$df 

modified_plots <- lapply(p_m2$plots, function(x) {
  x + theme(axis.title.x = element_text(vjust = 6.5, hjust = 1))
})

ggarrange(modified_plots$df, modified_plots$sse, modified_plots$ocv, modified_plots$gcv, ncol=2, nrow=2, common.legend = TRUE, legend="bottom")

p_m2$summary_panel
```

With $\text{PEN}_{m-2}(x)$, according to the GCV measure, the minimizing $m$ turns out to be
`r p_m2$opt_order$degree` with $\lambda=$ `r round(10^p_m2$opt_order$log_lambda,2)`
for which $\text{df}(\lambda)=$ `r round(p_m2$opt_order$df,2)`.
<!-- (Figure \ref{nchi_selection}, bottom-right panel). -->
<!-- <!-- (Figure \ref{nM_selection}, bottom-right panel). --> 

### PEN$_{2}$

We repeat the smoothing by posing
$\text{PEN}_{2}(x) =\int[D^{2}x(s)]^2 ds$ for each $m>3$ leaving
unaltered the other penalizations (Note: R routine `smooth.basis` gives
error posing `Lfd=1` with $m=2$ and `Lfd=2` with $m=3$, that is
penalizing a derivative consisting of a step function.) 
<!-- Figure \ref{nchi_selection2}). -->


```{r selection_2, fig.show='hold', fig.cap=paste("Smoothing selection using PEN$_{2}$ (",normtytxt," normalized data).",sep=""), fig.pos="p", w=4, h=3, echo=track}

p_2 <- smoothingSelection(
                      datnchi,
                      lambda_seq = NULL,
                      degrees = NULL,
                      penalty_type = "2",
                      plot = TRUE,
                      verbose = TRUE)
```

```{r optpar_2, dependson="selection_2", echo=track}
p_2$summary_optimal
p_2$optimal_gcv
p_2$optimal_ocv
p_2$opt_order

p_2$plots$df 

modified_plots <- lapply(p_2$plots, function(x) {
  x + theme(axis.title.x = element_text(vjust = 6.5, hjust = 1))
})

ggarrange(modified_plots$df, modified_plots$sse, modified_plots$ocv, modified_plots$gcv, ncol=2, nrow=2, common.legend = TRUE, legend="bottom")

p_2$summary_panel
```

With $\text{PEN}_{2}(x)$, according to the GCV measure, the minimizing $m$ results `r p_2$opt_order$degree` with $\lambda=$
`r round(10^p_2$opt_order$log_lambda,2)` for which $\text{df}(\lambda)=$
`r round(p_2$opt_order$df,2)`.
<!-- (Figure \ref{nchi_selection2}, bottom-right panel). -->

Lastly, we try with $\text{PEN}_{1}(x) =\int[Dx(s)]^2 ds$ for each $m>2$
and $\text{PEN}_{0}(x) =\int[D^0x(s)]^2 ds=\int[x(s)]^2 ds$ for every $m$. 
<!-- (Figures \ref{nchi_selection1} and \ref{nchi_selection0}). -->

### PEN$_{1}$

```{r selection_1, fig.show='hold', fig.cap=paste("Smoothing selection using PEN$_{1}$ (",normtytxt," normalized data).",sep=""), fig.pos="p", w=4, h=3, echo=track}

p_1 <- smoothingSelection(
                      datnchi,
                      lambda_seq = NULL,
                      degrees = NULL,
                      penalty_type = "1",
                      plot = TRUE,
                      verbose = FALSE)
```

```{r optpar_1, dependson="selection_1", echo=track}
p_1$summary_optimal
p_1$optimal_gcv
p_1$optimal_ocv
p_1$opt_order

p_1$plots$df 

modified_plots <- lapply(p_1$plots, function(x) {
  x + theme(axis.title.x = element_text(vjust = 6.5, hjust = 1))
})

ggarrange(modified_plots$df, modified_plots$sse, modified_plots$ocv, modified_plots$gcv, ncol=2, nrow=2, common.legend = TRUE, legend="bottom")

p_1$summary_panel
```

### PEN$_{0}$


```{r selection_0, fig.show='hold', fig.cap=paste("Smoothing selection using PEN$_{0}$ (",normtytxt," normalized data).",sep=""), fig.pos="p", w=4, h=3, echo=track}

p_0 <- smoothingSelection(
                      datnchi,
                      lambda_seq = NULL,
                      degrees = NULL,
                      penalty_type = "0",
                      plot = TRUE,
                      verbose = FALSE)
```

```{r optpar_0, dependson="selection_0", echo=track}
p_0$summary_optimal %>% dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 6)))
p_0$optimal_gcv
p_0$optimal_ocv
p_0$opt_order

p_0$plots$df 

modified_plots <- lapply(p_0$plots, function(x) {
  x + theme(axis.title.x = element_text(vjust = 6.5, hjust = 1))
})
ggarrange(modified_plots$df, modified_plots$sse, modified_plots$ocv, modified_plots$gcv, ncol=2, nrow=2, common.legend = TRUE, legend="bottom")

p_0$summary_panel
```

In these last cases, the minimizing value is $m=$ `r p_1$opt_order$degree` and $m=$
`r p_0$opt_order$degree` respectively, (with $\text{Pen}_1$, $\lambda=$
`r round(10^p_1$opt_order$log_lambda,2)` for which $\text{df}(\lambda)=$
`r round(p_1$opt_order$df,2)`, with $\text{Pen}_0$, $\lambda=$
`r round(10^p_0$opt_order$log_lambda,2)` for which $\text{df}(\lambda)=$
`r round(p_0$opt_order$df,2)`).

### Optimal smoothing

```{r optimalsmoothing, fig.show='hold', fig.cap=paste("Optimal smoothing comparison for different roughness penalties (",normtytxt," normalized data).",sep=""), fig.pos="p", echo=track}

resSmoothing <- list("m-2" = p_m2,
                     "2"   = p_2,
                     "1"   = p_1,
                     "0"   = p_0)

opt_res <- optimalSmoothing(resSmoothing)

# Results
#old <- options(pillar.sigfig = 6)
opt_res$opt_summary %>% print   # optimal values for all penalties
#options(old)
opt_res$penalty_opt   # optimal penalty type
opt_res$m_opt         # optimal order
opt_res$llambda_opt   # optimal log10(lambda)
opt_res$top4_opt
opt_res$plots$lLambda # plots
opt_res$plots$df
opt_res$plots$sse
opt_res$plots$gcv
```

```{r optimalsmoothing_plot, dependson=c("optimalsmoothing"), fig.show='hold', fig.cap=paste("Optimal smoothing comparison plots for different roughness penalties (",normtytxt," normalized data).",sep=""), fig.pos="p", echo=track}
modified_plots <- lapply(opt_res$plots, function(x) {
  x + xlab("m") + theme(
    plot.title = element_blank(),
    axis.title.x = element_text(vjust = 6.5, hjust = 0)
  )
})

ggarrange(modified_plots$lLambda, modified_plots$df, modified_plots$sse, modified_plots$gcv, ncol=2, nrow=2, common.legend = TRUE, legend="bottom")
```

In conclusion, the minimum GCV is achieved by setting $m=$ `r opt_res$m_opt` with
$\text{PEN}_{`r opt_res$penalty_opt`}$ roughness penalty
(note that for the top-4 types of penalizations,  `r opt_res$top4_opt`, the gcvs are almost identical or very close).

Optimal smoothing is illustrated below.
<!-- (Figures \ref{nchi_opt_fd}-\ref{nchi_opt_fit}). -->
<!-- \ref{nchi_opt_fd}-\ref{nchi_opt_fit}). -->

```{r optimal_fd, w=8, h=4.65, fig.cap=paste("Fit of the optimal smoothing spline to ",normtytxt," normalized data.",sep=""), echo=track, eval=T}

plot_opt <- plotOptimalSmoothing(datnchi, opt_result = opt_res, nx=F)
plot_opt$plot
```


```{r optimal_fit, dependson=c("fdautils","optimal_fd"), w=4, h=4.25, out.width='65%', fig.cap=paste("A selection of fitted curves according to increasing RMS residual. Fit of the optimal smoothing spline to ",normtytxt," normalized data.",sep=""), fig.pos="p", echo=track,  message = F, eval=T}

plotOptFit <- plotOptimalFits(data=datnchi,
                              smoothed = plot_opt$smoothed,
                              #opt_res=opt_res,
                              n_curves = 9,
                              show_zone = TRUE,
                              graph = FALSE)
plotOptFit$combinedKeywordPlot #%>% suppressWarnings(print())
#> $singleKeywordPlot
# Best and worst fitted curves
plotOptFit$singleKeywordPlot$effect #%>% suppressWarnings(print())
plotOptFit$singleKeywordPlot$riski #%>% suppressWarnings(print())
```

```{r savefinal_b, echo=F, message=F, eval=raw}
# save the smooth values (on fdtime) for internal use
nam <- paste0("smooth", normty)
assign(nam, plot_opt$smooth_vals)
# over must be TRUE to update internal data, but if it is TRUE it cancels existing objects without warning
# over <- T
# smoothnchi <- get(nam)
# usethis::use_data(smoothnchi, internal = T, overwrite = over)
# Then, we may use the following to save it in sysdata.rda
# OR, I can save it in data-raw/ and load it when necessary
save(list=nam, file=paste0("../data-raw/",nam,".rda"))
# OR, I can save it together with the objects in sysdata.rda
usethis::use_data(datnchi, kw_list, smoothnchi, internal = T, overwrite = over)
```
