---
title: "Tutorial: EDA"
author: "matilde trevisani"
date: "`r Sys.Date()`"
output: 
 rmarkdown::html_vignette:
   toc: true
   toc_depth: 3
   dev: CairoPNG
   gallery: true
   thumbnails: true    # render figures as "miniatures"
   lightbox: true    # figures can be enlarged when clicked
   highlight: espresso    # textmate  breezedark  haddock  espress monochrome  pygments
   df_print: "tibble" # other options: paged, kable, tibble
   code_folding: show
vignette: >
  %\VignetteIndexEntry{Tutorial: EDA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r optcmd, echo=FALSE, cache=F}
draft <- T
track <- T
# figeval <- T
# res <- "markup"
# raw <- T  # load/save to data-raw (within installation generates error)
# over=T
```

```{r utils, echo=F}
nam_work <- "JPSP"
```



```{r setup, include=FALSE, cache=F}
knitr::opts_chunk$set(echo = track, fig.path = paste0(nam_work,'_fig/eda/'), collapse = TRUE, comment = "#>", cache = 2, cache.path=paste0(nam_work,'_cache/'), cache.comments = F, autodep = TRUE, cache.extra = knitr::rand_seed)
knitr::opts_chunk$set(fig.align = 'center', crop = TRUE) 
knitr::opts_chunk$set(dev = c('CairoPNG','CairoPDF')) 
knitr::opts_knit$set(aliases=c(h = 'fig.height', w = 'fig.width'), unnamed.chunk.label = "fig")
#knit_hooks$set(crop = hook_pdfcrop)
#opts_knit$set(global.par = TRUE)
```



```{r lib, include=FALSE, cache=F}
# # Already in DESCRIPTION the following libraries
library(magrittr) # %*%, divide_by
# library(dplyr) 
# library(ggplot2)
# library(reshape2) # melt
# library(gridExtra) # grid.arrange
library(ggpubr) # ggarrange
library(tibble) # as_tibble, rownames_to_column
# #library(tidyverse) # contains
# # dplyr, ggplot2, magrittr, tibble, and many others
library(cccc)
```


# Data

## Raw data

For illustration we consider the data extracted from the titles of *Journal of Personality and Social Psychology* articles from 1965 to 2021. They consist in a term-document matrix (TDM) with lexical units (words or multi-word expressions) by row and their frequencies (in the titles of articles published in each year) by column, and a corpus-level metadata table with the total number of words in the articles (`Dim`), the number of articles (`Nart`) and  the reference volume for each year.
```{r rawdata, echo=track}
# load raw data.
# tdm and corpus raw data can be read from extdata folder of cccc library
tdm_filename <- system.file("extdata", "tdm.csv", package = "cccc")
corpus_filename <- system.file("extdata", "corpus.csv", package = "cccc")
tdm <- read.csv2(tdm_filename)
corpus <- read.csv2(corpus_filename)  
tdm 
corpus 
```
Raw data must be in CSV or Excel format, with the TDM having keywords^[We call "keywords" all the lexical units represented in the TDM as they are obtained from a selection of the original textual units extracted from the corpus.] in the first column and yearly frequencies in subsequent columns, and the corpus metadata containing at least year, total number of tokens and number of documents per year (second and third column respectively, any additional metadata in the fourth column if present). Note that time-points defining the subcorpora (of the diachronic corpus) may be of any temporal scale (most frequently they are years).

Raw data are processed by the function `importData()` which standardizes the format, renames columns, cleans keywords, and computes the total frequency per keyword. Furtherly, it classifies terms into frequency zones and creates a column representing the frequency interval associated with each zone.
```{r cleanup, message=T, eval=T}
data <- importData(tdm_file = tdm_filename, corpus_file = corpus_filename)
tdm <- data$tdm
corpus_info <- data$corpus_info
tdm 
corpus_info
```
Other information returned by `importData()` are whether the TDM has been normalized (default is FALSE), which columns in the TDM refer to yearly frequencies, the unique frequency zones and other items for graphical use.
```{r info, echo=track}
names(data)
data$norm; data$year_cols
```

### Frequency class

Keyword total frequency is highly variable. For convenience, keywords are grouped by their frequency class.
The default zone classification method is statistical, based on quartiles or balanced classes. The zones and their associated frequency intervals are:
```{r zones}
data$zone; data$int_freq %>% noquote()
tdm$zone %>% table()
# check quartiles 
cut(tdm$tot_freq, breaks=quantile(tdm$tot_freq, probs = seq(0,1, by=0.25)), include.lowest = T) %>% table()
```

We can also switch to a classification method based on linguistic frequency zones, then obtaining:
```{r linguistic, message=T, eval=T}
data_ling <- importData(tdm_file = tdm_filename, corpus_file = corpus_filename, zone="ling", verbose = F)
tdm_ling <- data_ling$tdm
tdm_ling 
data_ling$zone
data_ling$int_freq %>% noquote()
```

Function `rowMassPlot`illustrates graphically the word total frequency and the zone partition by quartiles and by linguistic frequency zones (Figure \ref{fig:rowmass}).

```{r, eval = F}
rowMassPlot(data)
rowMassPlot(data_ling)
```
```{r rowmass, echo = F, out.width='80%', w = 6, h = 2.8, warning=track, fig.cap="Keyword total frequencies and zones based on quartiles (left) or on linguistic frequency zones (right)", fig.label="fig:rowmass", fig.show='hold', fig.pos="p"}
p1 <- rowMassPlot(data)
p2 <- rowMassPlot(data_ling)
ggarrange(p1, ggplot() + theme_void(), p2, nrow = 1, widths = c(1, 0.1, 1))
#labels = c("A", "", "B")
```



### Subcorpora dimension

We consider four different sizes for subcorpora dimension (in view of normalization by column): number of documents (`nDoc`), total number of word-tokens in documents (`dimCorpus`), sum of keyword frequencies (column sum) in the TDM (`Csum`), maximum keyword frequency (maximum column frequency) in the TDM (`Mcf`). The first two are in the corpus metadata, the latter two can be computed from the TDM. The four measures are in general strongly correlated (except for `Mcf`). They can be visualized graphically as bar plots or line plots (Figure \ref{fig:colmass}).

```{r colscale, echo=track}
ind_d <- data$year_cols
Csum <- tdm[,ind_d] %>% apply(2,sum,na.rm=T)
Mcf <- tdm[,ind_d] %>% apply(2,max,na.rm=T)
d_size <- corpus_info[,c(1,3,2)] %>% cbind(Csum, Mcf)
d_size 
d_size[,-1] %>% summary
d_size[,-c(1,5)] %>% apply(2,sum)
```

```{r rescaling}
# Identify a proper rescaling before making the plot of temporal dimensions.
sc <- c(1, 10, 10, 1)
apply(d_size[,-1], 2, max) %>% divide_by(sc)
ymax <- apply(d_size[,-1], 2, max) %>% divide_by(sc) %>% max %>% print
```

```{r colmass, echo=track, results=track, w=8, h=4, fig.cap="Subcorpora dimension: for each year, number of texts (dot-line), total number of word-tokens(/$10$) in texts, column sum(/$10$) in the TDM, maximum column frequency in the TDM", fig.label="fig:colmass", fig.pos="p", out.width='100.0%'}

colMassPlot(data, sc = c(1, 10, 10, 1), r = 1, textty = "text", themety = "light", size_b = 2.5, x_lab = "year")

```


### Temporal curves

The overall frequency trajectories (curves) are shown for the original data (Figure \ref{fig:curves}) and, below, also separately for each frequency class or zone (with a sample for each class) (Figure \ref{fig:facets}).

The plots are based on the long-format TDM stored in `data$tdm_long`, which has `year` as a group variable and `frequency` as the associated numerical variable.
```{r dat_l, include=track, eval = T}
tdm_l <- data$tdm_long 
tdm_l %>% head; tdm_l %>% tail
```

```{r curves, eval=T, echo=track, w=8, h=5, fig.cap="Keyword trajectories (original data)", fig.label="fig:curves", fig.pos="p", out.width='100.0%'}

curvePlot(data, r = 2, size_class = c(.2,.3,.4,.5))
```

```{r sample, message=T, include=T}
r <- 3
# seed <- Sys.time()
seed <- 2222
set.seed(seed)
kw <- tdm_l %>% aggregate(keyword ~ zone, ., function(x) sample(unique(x), r))
kw
```


```{r curvesEtctu, eval=T, dependson=c("sample"), echo=track, w=8, h=5, fig.cap="Keyword trajectories (original data)", fig.label="fig:curvesEtctu", fig.pos="p", out.width='100.0%'}

ctu_noun <-  c("person", "object", "instruct", "incident")

ctu_noun <- kw %>% pull(2) %>% as.data.frame() %>% pull(1)

curveCtuPlot(data, ctu_noun = ctu_noun, r = 2, size_class = NULL)
```

```{r example_with_frequency, eval=T, message = F}
library(dplyr)
tdm %>% group_by(zone) %>% arrange(desc(tot_freq)) %>% slice_head(n = 3) %>% ungroup() %>% select(zone, keyword)
```


```{r facets, echo=track, w=7.5, h=6, fig.cap="Keyword trajectories by frequency class (original data). Group data scale on y-axis.", fig.label="fig:facets", fig.pos="p", out.width='100.0%'}

facetPlot(data, keyword_selection = list(type="frequency", n=3, kw.list=NULL),
r = 4, scales = "free", leg = TRUE, size_class = NULL)
```


# Normalization


```{r normalize, include=draft}

options(digits=3)

cat("\n###### normalization by column ######\n
\n## subcorpus total #tokens (dimCorpus) ##\n")

datnc <- normalization(data, normty = "nc", 1000)

# datnc[,1:6] %>% head 
# datnc[,-(1:2)] %>% max
# # datnc[,-(1:2)] %>% apply(2,max)


cat("\n###### normalization by row and col ######\n
\n## chisquare-like ##\n")

datnchi <- normalization(data, normty = "nchi", 1)

# datnchi[,1:6] %>% head 
# datnchi[,1:6] %>% tail 
# datnchi[,-(1:2)] %>% max
# # datnchi[,-(1:2)] %>% apply(2,max)

cat("\n##  Max row freq and subcorpus total #tokens  ##\n\n")

datnM <- normalization(data, normty = "nM", 1000)

# datnM[,1:6] %>% head
# # datnM[,-(1:2)] %>% apply(1,max)
# # datnM[,-(1:2)] %>% apply(1,min)
# #datnM[,-(1:2)] %>% apply(2,max)


cat("\n## MinMax row freq and subcorpus total #tokens  ##\n")

datnmM <- normalization(data, normty = "nmM", 1000)

# datnmM[,1:6] %>% head 
# # datnmM[,-(1:2)] %>% apply(1,max)
# # datnmM[,-(1:2)] %>% apply(1,min)
# #datnmM[,-(1:2)] %>% apply(2,max)
# #datnmM[,-(1:2)] %>% apply(2,min)

cat("\n## Non-linear row freq and subcorpus total #tokens ##\n")

cat("\nNon-linear transformation:\n Difference from the minimum weights more when the distribution is positively asymmetric (p>0.5), viceversa, difference from the maximum weights more when the distribution is negatively asymmetric(p<0.5)\n")


datnnl <- normalization(data, normty = "nnl", 1000, nnlty="M")
# p_asy <- datnnl$p_asy
# datnnl <- datnnl$datn
# 
# # # kw 379 has VL=1, p=1, then 0/0 when x=min
# # datnnl[379,]
# # datnnl[379,2+which(apply(datnnl[379,-(1:2)], 1, is.nan))] <- 0
# # datnnl[379,]
# 
# # check any nonfinite value
# apply(datnnl[,-(1:2)], 1, function(x){any(is.na(x))}) %>% sum
# apply(datnnl[,-(1:2)], 1, function(x){any(is.infinite(x))}) %>% sum
# apply(datnnl[,-(1:2)], 1, function(x){any(is.nan(x))}) %>% sum
# 
# datnnl[,1:6] %>% head
# datnnl[,-(1:2)] %>% apply(1,max)
# datnnl[,-(1:2)] %>% apply(1,min)
# #datnmM[,-(1:2)] %>% apply(2,max)
# #datnmM[,-(1:2)] %>% apply(2,min)
# 
# # cat("\nComparing MinMax to non-linear norm:\n")
# # cbind(datnmM %>% filter(keyword%in%c("metro","prendere")) %>% t(),
# # datnnl %>% filter(keyword%in%c("metro","prendere")) %>% t())[c(1:5, (n_d-4):n_d),]

```

```{r}
names(datnc)
datnc$norm; datnc$normty
datnc$tdm %>% head
datnc$tdm_long %>% head
```


```{r asy_eda, eval=T}
options(digits=3)

names(datnnl)
temp <- datnnl$tdm[,c(1,4)] %>% cbind(datnnl$p_asy)
# table(temp$zone, datnnl$p_asy>.5)
# # temp %>% filter(p_asy<.5)
# temp %>%
#   group_by(zone) %>%
#   summarise(min=min(p_asy), p25=quantile(p_asy,.25),me=median(p_asy),p75=quantile(p_asy,.75),max=max(p_asy), mean=mean(p_asy))

```


