---
title: "Tutorial: Exploratory Data Analysis (EDA)"
author: "matilde trevisani"
date: "`r Sys.Date()`"
output: 
 rmarkdown::html_vignette:
   toc: true
   toc_depth: 3
   fig_width: 8
   fig_height: 5
   # highlight: espresso    # textmate  breezedark  haddock  espress monochrome  pygments
   df_print: "tibble" # other options: paged, kable, tibble
vignette: >
  %\VignetteIndexEntry{Tutorial: Exploratory Data Analysis (EDA)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r optcmd, echo=FALSE, cache=F}
draft <- T
track <- T
figeval <- T
raw <- T # T when locally knitting, set to F when creating vignettes within installation
over <- T # overwrite cached objects
```

```{r utils, echo=F}
nam_work <- "JPSP"
```


```{r setup, include=FALSE, cache=F}
knitr::opts_chunk$set(
  echo = track, 
  collapse = TRUE, 
  comment = "#>", 
  fig.path = paste0(nam_work,'_fig/eda/'), 
  cache = 2, 
  cache.path=paste0(nam_work,'_cache/'), 
  cache.comments = F, 
  autodep = TRUE, 
  cache.extra = knitr::rand_seed,
  fig.align = 'center',
  #crop = TRUE,
  fig.width = 8,
  fig.height = 5,
  out.width = '100%',
  warning = track,
  message = track
  )

# Cairo opzionale per figure di alta qualità
if (requireNamespace("Cairo", quietly = TRUE)) {
  knitr::opts_chunk$set(dev = 'png', dev.args = list(type = "cairo"))
}

knitr::opts_knit$set(aliases=c(h = 'fig.height', w = 'fig.width'), unnamed.chunk.label = "fig")
```



```{r packages, include=FALSE, cache=F}
# Main package
library(cccc)

# Data manipulation
library(dplyr)
library(tibble) # as_tibble, rownames_to_column

# Visualization
library(ggplot2)
#library(patchwork)

# Interactive tables
library(DT)

library(magrittr) # %*%, non serve più per divide_by
library(ggpubr) # get_legend, può servire ancora per ggarrange
```


# Data

## Raw data

For illustration we consider the data extracted from the titles of *Journal of Personality and Social Psychology* articles from 1965 to 2021. They consist in a term-document matrix (TDM) with lexical units (words or multi-word expressions) by row and their frequencies (in the titles of articles published in each year) by column, and a corpus-level metadata table with the total number of words in the articles (`Dim`), the number of articles (`Nart`) and  the reference volume for each year.
```{r load-data, echo=track}
# Load raw data from package
tdm_filename <- system.file("extdata", "tdm.csv", package = "cccc")
corpus_filename <- system.file("extdata", "corpus.csv", package = "cccc")

# Check files exist
if (tdm_filename == "" || corpus_filename == "") {
  stop("Example data files not found")
}

tdm <- read.csv2(tdm_filename)
corpus <- read.csv2(corpus_filename)  
```

```{r show-raw-data, results=draft}
datatable(
  tdm,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    scrollY = "400px",
    dom = 'frtip'
  ),
  caption = "Term-Document Matrix (raw data)"
)
datatable(
  corpus,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Corpus Information"
)
```

Raw data must be in CSV or Excel format, with the TDM having keywords^[We call "keywords" all the lexical units represented in the TDM as they are obtained from a selection of the original textual units extracted from the corpus.] in the first column and yearly frequencies in subsequent columns, and the corpus metadata containing at least year, total number of tokens and number of documents per year (second and third column respectively, any additional metadata in the fourth column if present). Note that time-points defining the subcorpora (of the diachronic corpus) may be of any temporal scale (most frequently they are years).

Raw data are processed by the function `importData()` which standardizes the format, renames columns, cleans keywords, and computes the total frequency per keyword. Furtherly, it classifies terms into frequency zones and creates a column representing the frequency interval associated with each zone.

```{r import-data}
data <- importData(
  tdm_file = tdm_filename, 
  corpus_file = corpus_filename,
  sep_tdm = ";",
  sep_corpus_info = ";",
  zone = "stat",
  verbose = FALSE
  )

tdm <- data$tdm
corpus_info <- data$corpus_info
```
```{r show-imported-data, results=draft}
datatable(
  tdm,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    scrollY = "400px",
    dom = 'frtip'
  ),
  caption = "Term-Document Matrix (processed)"
)
datatable(
  corpus_info,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Corpus Information (processed)"
)
```

Other information returned by `importData()` are whether the TDM has been normalized (default is FALSE), which columns in the TDM refer to yearly frequencies, the unique frequency zones and other items for graphical use.

```{r data-info, echo=track}
names(data)
data$norm
data$year_cols
```

### Frequency class

Keyword total frequency is highly variable. For convenience, keywords are grouped by their frequency class.
The default zone classification method is statistical, based on quartiles or balanced classes. The zones and their associated frequency intervals are:
```{r zones}
data$zone
noquote(data$int_freq)
table(tdm$zone)

# check quartiles 
cut(
  tdm$tot_freq, 
  breaks=quantile(tdm$tot_freq, probs = seq(0,1, by=0.25)),
  include.lowest = T
  ) %>% table()
```

We can also switch to a classification method based on linguistic frequency zones:

```{r linguistic-zones}
data_ling <- importData(
  tdm_file = tdm_filename, 
  corpus_file = corpus_filename, 
  zone="ling", 
  verbose = F
  )

tdm_ling <- data_ling$tdm
data_ling$zone
noquote(data_ling$int_freq)
```

```{r show-linguistic-tdm, results=draft}
datatable(
  tdm_ling,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    scrollY = "400px",
    dom = 'frtip'
  ),
  caption = "Term-Document Matrix with linguistic frequency zones"
)
```

Function `rowMassPlot`illustrates graphically the word total frequency and the zone partition by quartiles and by linguistic frequency zones. 
<!-- (Figure \ref{fig:rowmass}). -->

```{r, eval = F}
rowMassPlot(data)
rowMassPlot(data_ling)
```
```{r rowmass-plot, echo = F, out.width='80%', w = 6, h = 2.8, warning=track, fig.cap="Keyword total frequencies and zones based on quartiles (left) or on linguistic frequency zones (right)", fig.label="fig:rowmass", fig.show='hold', fig.pos="p"}
p1 <- rowMassPlot(data)
p2 <- rowMassPlot(data_ling)
ggarrange(p1, ggplot() + theme_void(), p2, nrow = 1, widths = c(1, 0.1, 1))
# #labels = c("A", "", "B")
#p1 + plot_spacer() + p2 + plot_layout(widths = c(1, 0.1, 1))
```



### Subcorpora dimension

We consider four different sizes for subcorpora dimension (in view of normalization by column): number of documents (`nDoc`), total number of word-tokens in documents (`dimCorpus`), sum of keyword frequencies (column sum) in the TDM (`Csum`), maximum keyword frequency (maximum column frequency) in the TDM (`Mcf`). The first two are in the corpus metadata, the latter two can be computed from the TDM. The four measures are in general strongly correlated (except for `Mcf`). They can be visualized graphically as bar plots or line plots. 
<!-- (Figure \ref{fig:colmass}). -->

```{r subcorpora-size, echo=track, results=draft}
ind_d <- data$year_cols
Csum <- tdm[,ind_d] %>% apply(2,sum,na.rm=T)
Mcf <- tdm[,ind_d] %>% apply(2,max,na.rm=T)
d_size <- corpus_info[,c(1,3,2)] %>% cbind(Csum, Mcf)

datatable(
  d_size,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Subcorpora dimensions by year"
)

summary(d_size[,-1])
apply(d_size[,-c(1,5)], 2,sum)
```

Identify proper rescaling before plotting:

```{r rescaling}
# Identify a proper rescaling before making the plot of temporal dimensions.
sc <- c(1, 10, 10, 1)
apply(d_size[,-1], 2, max) / sc
ymax <- max(apply(d_size[,-1], 2, max) / sc) 
ymax
```

```{r colmass, echo=track, results=track, w=8, h=4, fig.cap="Subcorpora dimension: for each year, number of texts (dot-line), total number of word-tokens(/10) in texts, column sum(/10) in the TDM, maximum column frequency in the TDM", fig.label="fig:colmass", fig.pos="p"}

colMassPlot(
  data, 
  sc = c(1, 10, 10, 1), 
  r = 1, 
  textty = "text", 
  themety = "light", 
  size_b = 2.5, 
  x_lab = "year"
  )

```


### Temporal curves

The overall frequency trajectories (curves) are shown for the original data and, below, also separately for each frequency class or zone (with a sample for each class). 
<!-- (Figures \ref{fig:curves} and \ref{fig:facets}). -->

The plots are based on the long-format TDM stored in `data$tdm_long`, which has `year` as a group variable and `frequency` as the associated numerical variable.

```{r tdm_long, include=track}
tdm_l <- data$tdm_long

datatable(
  tdm_l,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    scrollY = "400px",
    dom = 'frtip'
  ),
  caption = "Term-Document Matrix in long format"
)
```

```{r curve-plot, eval=T, echo=track, w=8, h=5, fig.cap="Keyword trajectories (original data)", fig.label="fig:curves", fig.pos="p"}

curvePlot(data, r = 2, size_class = c(0.1, 0.2, 0.35, 0.35))
```

A sample of keywords (three for each zone) can be extracted, randomly, as follows:

```{r sample-keywords, include=T}
r <- 3
set.seed(8)
kw <- tdm_l %>% group_by(zone) %>% summarise(keyword = list(sample(unique(keyword), r))) %>% ungroup()
kw

kw <- tdm_l %>% aggregate(keyword ~ zone, ., function(x) sample(unique(x), r))
kw
```

Hence, we can highlight a sample of zone-specific keywords (randomly generated as above or manually imputed) over the curves.
```{r curves-ctu-plot, dependson=c("sample-keywords"), echo=track, w=8, h=5, fig.cap="Keyword trajectories with example keywords highlighted", fig.label="fig:curvesEtctu", fig.pos="p"}

#ctu_noun <- kw %>% pull(2) %>% as.data.frame() %>% pull(1)
ctu_noun <-  c("prejudic", "contact", "correspond", "decod")

curveCtuPlot(data, ctu_noun = ctu_noun, r = 2)
```

Top Keywords (i.e., with highest total frequency) per zone can also be selected.
```{r top-Keywords}
# select top 3 keywords by total frequency per zone
top_kw <- tdm %>% 
  group_by(zone) %>% 
  arrange(desc(tot_freq)) %>% 
  slice_head(n = 3) %>% 
  ungroup() %>% 
  select(zone, keyword, tot_freq)

datatable(
  top_kw,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    dom = 't'
  ),
  caption = "Top 3 keywords by frequency zone"
)
```
This kind of selection corresponds to the `type = frequency` in the function `facetPlot()`, as shown below.

```{r facet-plot, echo=track, w=7.5, h=6, fig.cap="Keyword trajectories by frequency class (original data). Group data scale on y-axis.", fig.label="fig:facets", fig.pos="p"}

facetPlot(
  data, 
  keyword_selection = list(type="frequency", n=3),
  r = 4, 
  scales = "free",
  leg = TRUE, 
  size_class = NULL
  )
```
Let consider a list of keywords user-defined which can be imputed by setting `type = list` (as opposed to `random` or, like above, `frequency`).

```{r user_kw, w=7.5, h=6, fig.cap="Keyword trajectories by frequency class (original data). Group data scale on y-axis.", fig.label="fig:facets.list", fig.pos="p"}

kw_list <- c("prejudic", "problem", "trust", "contact", "order", "account", "correspond", "space", "group decis make", "decod", "physiolog respons", "smoke")

facetPlot(data, keyword_selection = list(type="list", n=3, kw.list=kw_list), r = 4, scales = "free")
```


# Normalization

There are different types of normalization that can be applied to raw frequencies in order to reduce the influence of subcorpus sizes and keyword overall frequencies on the analysis of temporal trajectories. The function `normalization()` implements five types of normalization: by column only (`nc`) as well as by row and column (`nchi`, `nM`, `nmM`, `nnl`). See the documentation for details.
```{r normalize, echo=-(c(1:4, 6:7, 12:15, 19, 21, 24:26, 31:32, 36:37, 42:45, 49:50, 54:70)), include=draft}

options(digits=3)

cat("###### original data ######\n")
data$tdm_long %>% group_by(zone) %>% summarize(max=max(freq), min=min(freq), mean=mean(freq), median=median(freq), q25=quantile(freq,.25), q75=quantile(freq,.75))

cat("\n###### normalization by column ######\n
## subcorpus total #tokens (dimCorpus) ##")

datnc <- normalization(data, normty = "nc", 1000)

datnc$norm; datnc$normty
# datnc$tdm[, 1:9] %>% head
datnc$tdm_long %>% group_by(zone) %>% summarize(max=max(freq), min=min(freq), mean=mean(freq), median=median(freq), q25=quantile(freq,.25), q75=quantile(freq,.75))

cat("\n###### normalization by row and col ######\n
## chisquare-like ##")

datnchi <- normalization(data, normty = "nchi", 1)

datnchi$tdm_long %>% group_by(zone) %>% summarize(max=max(freq), min=min(freq), mean=mean(freq), median=median(freq), q25=quantile(freq,.25), q75=quantile(freq,.75))

cat("\n##  Max row freq and subcorpus total #tokens  ##\n\n")

datnM <- normalization(data, normty = "nM", 1000)

# datnM$tdm[, 1:9] %>% head
datnM$tdm_long %>% group_by(zone) %>% summarize(max=max(freq), min=min(freq), mean=mean(freq), median=median(freq), q25=quantile(freq,.25), q75=quantile(freq,.75))

ind_d <- data$year_cols
datnM$tdm[, ind_d] %>% apply(1,max) %>% summary
datnM$tdm[, ind_d] %>% apply(1,min) %>% summary

cat("\n## MinMax row freq and subcorpus total #tokens  ##\n")

datnmM <- normalization(data, normty = "nmM", 1000)

# datnmM$tdm[, 1:9] %>% head
datnmM$tdm_long %>% group_by(zone) %>% summarize(max=max(freq), min=min(freq), mean=mean(freq), median=median(freq), q25=quantile(freq,.25), q75=quantile(freq,.75))

datnmM$tdm[, ind_d] %>% apply(1,max) %>% summary
datnmM$tdm[, ind_d] %>% apply(1,min) %>% summary


cat("\n## Non-linear row freq and subcorpus total #tokens ##\n")

cat("Non-linear transformation:\n Difference from the minimum weights more when the distribution is positively asymmetric (p>0.5), viceversa, difference from the maximum weights more when the distribution is negatively asymmetric(p<0.5)")

datnnl <- normalization(data, normty = "nnl", 1000, nnlty="M")
p_asy <- datnnl$p_asy

datnnl$tdm_long %>% group_by(zone) %>% summarize(max=max(freq), min=min(freq), mean=mean(freq), median=median(freq), q25=quantile(freq,.25), q75=quantile(freq,.75))

datnnl$tdm[, ind_d] %>% apply(1,max) %>% summary
datnnl$tdm[, ind_d] %>% apply(1,min) %>% summary
# datnmM[, ind_d] %>% apply(2,max)
# datnmM[, ind_d] %>% apply(2,min)

# Non-linear norm with variance-based asymmetry index
# datnnl <- normalization(data, normty = "nnl", 1000, nnlty="V")
# p_asy <- datnnl$p_asy
# 
# # # kw 379 has VL=1, p=1, then 0/0 when x=min
# # datnnl[379,]
# # datnnl[379,2+which(apply(datnnl[379,-(1:2)], 1, is.nan))] <- 0
# # datnnl[379,]
# 
# # check any nonfinite value
# apply(datnnl$tdm[, ind_d], 1, function(x){any(is.na(x))}) %>% sum
# apply(datnnl$tdm[, ind_d], 1, function(x){any(is.infinite(x))}) %>% sum
# apply(datnnl$tdm[, ind_d], 1, function(x){any(is.nan(x))}) %>% sum

```
Some exploratory analysis on the asymmetry index `p_asy` used in the non-linear normalization and on the effect on keyword frequencies is shown below.
```{r asy_eda, echo = -c(1:2), eval=T}
options(digits=3)

temp <- datnnl$tdm %>% cbind(p_asy=datnnl$p_asy) 
# How many keywords have p_asy > 0.5
temp %$% table(zone, p_asy>.5)
# Summary statistics of p_asy by zone
temp %>%
  group_by(zone) %>%
  summarise(min=min(p_asy), p25=quantile(p_asy,.25),me=median(p_asy),p75=quantile(p_asy,.75),max=max(p_asy), mean=mean(p_asy))

cat("\nComparing MinMax to non-linear norm:\n")
kw_compare <- c("pain", "mental health")
temp %>% select(keyword, zone, p_asy) %>% filter(keyword%in%kw_compare)
rbind(datnmM$tdm %>% filter(keyword%in%kw_compare), datnnl$tdm %>% filter(keyword%in%kw_compare))[,c(1, 4:ind_d[5], last(ind_d)-(8:0))] %>% t() %>% noquote()


```

## Normalization by column

Normalization by column (subcorpus total \#tokens) (Figure \ref{fig:curvesnc}-\ref{fig:facetsnc})


```{r curvesnc, echo=track, w=8, h=5, fig.cap="Keyword trajectories (data normalized by column, subcorpus dimension)", fig.label="fig:curvesnc", fig.pos="p", out.width='100.0%', eval=draft}

curvePlot(datnc, r = 2, y_lab = "keyword (column normalized) frequency")

```

```{r curvesEtctunc, echo=track, results=draft, w=8, h=5, fig.cap="Keyword trajectories (data normalized by column, subcorpus dimension)", fig.label="fig:curvesEtctunc", fig.pos="p", out.width='100.0%', eval=draft}

curveCtuPlot(datnc, ctu_noun, r = 2, y_lab = "keyword (column normalized) frequency")

```

```{r facetsnc, echo=track, w=7.5, h=6, fig.cap="Keyword trajectories by frequency class (data normalized by column)", fig.label="fig:facetsnc", fig.pos="p", out.width='100.0%', eval=T}

facetPlot(datnc, keyword_selection = list(type="list", n=3, kw.list=kw_list), r = 4, scales = "free", y_lab = "keyword (column normalized) frequency")

```

## Chisquare-like (double) normalization 

Double normalization (chisquare-like) (Figure \ref{fig:curvesnchi}-\ref{fig:facetsnchi})


```{r curvesnchi, echo=track, w=8, h=5, fig.cap="Keyword trajectories (data doubly normalized, chisquare-like)", fig.label="fig:curvesnchi", fig.pos="p", out.width='100.0%', eval=figeval}

curvePlot(datnchi, r = 2, y_lab = "keyword (doubly normalized) frequency (X2-like)")

```

```{r curvesEtctunchi, echo=track, w=8, h=5, fig.cap="Keyword trajectories (data doubly normalized, chisquare-like)", fig.label="fig:curvesEtctunchi", fig.pos="p", out.width='100.0%', eval=figeval}

curveCtuPlot(datnchi, ctu_noun, r = 2, y_lab = "keyword (doubly normalized) frequency (X2-like)")

```

```{r facetsnchi, echo=track,  w=7.5, h=11, fig.cap="Keyword trajectories by frequency size class (data doubly normalized, chisquare-like)", fig.label="fig:facetsnchi", fig.pos="p", out.width='100.0%', eval=figeval}

p1 <- facetPlot(datnchi, keyword_selection = list(type="list", n=3, kw.list=kw_list), r = 4, leg = F, scales = "fixed", y_lab = "keyword (doubly normalized) frequency (X2-like)") + labs(x = NULL)
p2 <- facetPlot(datnchi, keyword_selection = list(type="list", n=3, kw.list=kw_list), r = 4,  scales = "free", y_lab = "keyword (doubly normalized) frequency (X2-like)")
legend <- get_legend(p2)
p2 <- p2 + theme(legend.position="none")
ggarrange(p1, p2, legend, nrow=3, heights=c(3.95, 4, .75))

```

## Word maximum frequency (double) normalization

Double normalization (by maximum row frequency) (Figure \ref{fig:curvesnM}-\ref{fig:facetsnM})


```{r curvesnM, echo=track, w=8, h=5, fig.cap="Keyword trajectories (data doubly normalized, maximum row frequency)", fig.label="fig:curvesnM",fig.pos="p", out.width='100.0%', eval=F}

curvePlot(datnM, r = 2, size_class = rep(.1,4), y_lab = "keyword (doubly normalized) frequency (Max row freq)")
```


```{r curvesEtctunM, echo=track, results=draft, w=8, h=5, fig.cap="Keyword trajectories (data doubly normalized, maximum row frequency)", fig.label="fig:curvesEtctunM", fig.pos="p", out.width='100.0%', eval=F}

curveCtuPlot(datnM, ctu_noun, r = 2, size_class = rep(.1,4), y_lab = "keyword (doubly normalized) frequency (Max row freq)")
```


```{r facetsnM, echo=track, w=7.5, h=6, fig.cap="Keyword trajectories by frequency size class (data doubly normalized, maximum row frequency)", fig.label="fig:facetsnM", fig.pos="p", out.width='100.0%', eval=F}

facetPlot(datnM, keyword_selection = list(type="list", n=3, kw.list=kw_list), r = 4, size_class = rep(.1,4), scales = "free", y_lab = "keyword (doubly normalized) frequency (Max row freq)")
```


## Minmax (double) normalization

Double normalization (by row minmax) (Figure \ref{fig:curvesnmM}-\ref{fig:facetsnmM})

```{r curvesnmM, echo=track, w=8, h=5, fig.cap="Keyword trajectories (data doubly normalized, row minmax)", fig.label="fig:curvesnmM",fig.pos="p", out.width='100.0%', eval=figeval}

curvePlot(datnmM, r = 2, size_class = rep(.1,4), y_lab = "keyword (doubly normalized) frequency (MinMax)")
```


```{r curvesEtctunmM, echo=track, results=draft, w=8, h=5, fig.cap="Keyword trajectories (data doubly normalized, row minmax)", fig.label="fig:curvesEtctunmM", fig.pos="p", out.width='100.0%', eval=figeval}

curveCtuPlot(datnmM, ctu_noun, r = 2, size_class = rep(.1,4), y_lab = "keyword (doubly normalized) frequency (MinMax)")
```


```{r facetsnmM, echo=track, w=7.5, h=6, fig.cap="Keyword trajectories by frequency size class (data doubly normalized, row minmax)", fig.label="fig:facetsnmM", fig.pos="p", out.width='100.0%', eval=figeval}

facetPlot(datnmM, keyword_selection = list(type="list", n=3, kw.list=kw_list), r = 4, size_class = rep(.1,4), scales = "free", y_lab = "keyword (doubly normalized) frequency (MinMax)")
```


## Non-linear (double) normalization

Double normalization (by nonlinear row frequency normalization) (Figure \ref{fig:curvesnnl}-\ref{fig:facetsnnl})

```{r curvesnnl, dependson=c("graphpar2", "datnnl_l"), echo=track, w=8, h=5, fig.cap="Keyword trajectories (data doubly normalized, non-linear M-based)", fig.label="fig:curvesnnl",fig.pos="p", out.width='100.0%', eval=T}

curvePlot(datnnl, r = 2, size_class = rep(.1,4), y_lab = "keyword (doubly normalized) frequency (non-linear M-based)")
```

What about the `p_asy` asymmetry index of the example keywords?

```{r}
temp %>% select(keyword, zone, p_asy) %>% filter(keyword%in%kw_list)
```


```{r curvesEtctunnl, echo=track, results=draft, w=8, h=5, fig.cap="Keyword trajectories (data doubly normalized, non-linear M-based)", fig.label="fig:curvesEtctunnl", fig.pos="p", out.width='100.0%', eval=T}

curveCtuPlot(datnnl, ctu_noun, r = 2, size_class = rep(.1,4), y_lab = "keyword (doubly normalized) frequency (non-linear M-based)")
```


```{r facetsnnl, echo=track, w=7.5, h=6, fig.cap="Keyword trajectories by frequency size class (data doubly normalized, non-linear M-based)", fig.label="fig:facetsnnl", fig.pos="p", out.width='100.0%', eval=T}

facetPlot(datnnl, keyword_selection = list(type="list", n=3, kw.list=kw_list), r = 4, size_class = rep(.1,4), scales = "free", y_lab = "keyword (doubly normalized) frequency (non-linear M-based)")
```


```{r savefinal, echo=F, message=F, eval=raw}
usethis::use_data(datnchi, kw_list, internal = T, overwrite = over)
```


