---
title: "Tutorial: Exploratory Data Analysis EDA"
author: "matilde trevisani"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    fig_width: 8
    fig_height: 5
vignette: >
  %\VignetteIndexEntry{Tutorial: Exploratory Data Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.align = 'center',
  fig.width = 8,
  fig.height = 5,
  out.width = '100%',
  warning = FALSE,
  message = FALSE
)

# Cairo opzionale per figure di alta qualità
if (requireNamespace("Cairo", quietly = TRUE)) {
  knitr::opts_chunk$set(dev = 'png', dev.args = list(type = "cairo"))
}
```

```{r packages, include=FALSE}
# Main package
library(cccc)

# Data manipulation
library(dplyr)
library(tibble)

# Visualization
library(ggplot2)
library(patchwork)

# Interactive tables
library(DT)
```

# Data

## Raw data

For illustration we consider the data extracted from the titles of *Journal of Personality and Social Psychology* articles from 1965 to 2021. They consist in a term-document matrix (TDM) with lexical units (words or multi-word expressions) by row and their frequencies (in the titles of articles published in each year) by column, and a corpus-level metadata table with the total number of words in the articles (`Dim`), the number of articles (`Nart`) and  the reference volume for each year.

```{r load-data}
# Load raw data from package
tdm_filename <- system.file("extdata", "tdm.csv", package = "cccc")
corpus_filename <- system.file("extdata", "corpus.csv", package = "cccc")

# Check files exist
if (tdm_filename == "" || corpus_filename == "") {
  stop("Example data files not found")
}

tdm <- read.csv2(tdm_filename)
corpus <- read.csv2(corpus_filename)
```

```{r show-raw-data}
datatable(
  tdm,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    scrollY = "400px",
    dom = 'frtip'
  ),
  caption = "Term-Document Matrix (raw data)"
)

datatable(
  corpus,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Corpus Information"
)
```

Raw data must be in CSV or Excel format, with the TDM having keywords^[We call "keywords" all the lexical units represented in the TDM as they are obtained from a selection of the original textual units extracted from the corpus.] in the first column and yearly frequencies in subsequent columns, and the corpus metadata containing at least year, total number of tokens and number of documents per year (second and third column respectively, any additional metadata in the fourth column if present). Note that time-points defining the subcorpora (of the diachronic corpus) may be of any temporal scale (most frequently they are years).

Raw data are processed by the function `importData()` which standardizes the format, renames columns, cleans keywords, and computes the total frequency per keyword. Furthermore, it classifies terms into frequency zones and creates a column representing the frequency interval associated with each zone.

```{r import-data}
data <- importData(
  tdm_file = tdm_filename, 
  corpus_file = corpus_filename,
  sep_tdm = ";",
  sep_corpus_info = ";",
  zone = "stat",
  verbose = FALSE
)

tdm <- data$tdm
corpus_info <- data$corpus_info
```

```{r show-imported-data}
datatable(
  tdm,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    scrollY = "400px",
    dom = 'frtip'
  ),
  caption = "Term-Document Matrix (processed)"
)

datatable(
  corpus_info,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Corpus Information (processed)"
)
```

Other information returned by `importData()` are whether the TDM has been normalized (default is FALSE), which columns in the TDM refer to yearly frequencies, the unique frequency zones and other items for graphical use.

```{r data-info}
names(data)
data$norm
data$year_cols
```

### Frequency class

Keyword total frequency is highly variable. For convenience, keywords are grouped by their frequency class.
The default zone classification method is statistical, based on quartiles or balanced classes. The zones and their associated frequency intervals are:

```{r zones}
data$zone
noquote(data$int_freq)
table(data$tdm$zone)

# Check quartiles
cut(
  tdm$tot_freq, 
  breaks = quantile(tdm$tot_freq, probs = seq(0, 1, by = 0.25)), 
  include.lowest = TRUE
) %>% table()
```

We can also switch to a classification method based on linguistic frequency zones:

```{r linguistic-zones}
data_ling <- importData(
  tdm_file = tdm_filename, 
  corpus_file = corpus_filename,
  sep_tdm = ";",
  sep_corpus_info = ";",
  zone = "ling", 
  verbose = FALSE
)

tdm_ling <- data_ling$tdm
data_ling$zone
noquote(data_ling$int_freq)
```

```{r show-linguistic-tdm}
datatable(
  tdm_ling,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    scrollY = "400px",
    dom = 'frtip'
  ),
  caption = "Term-Document Matrix with linguistic frequency zones"
)
```

Function `rowMassPlot` illustrates graphically the word total frequency and the zone partition by quartiles and by linguistic frequency zones.

```{r rowmass-plot, fig.width=10, fig.height=4, fig.cap="Keyword total frequencies and zones based on quartiles (left) or on linguistic frequency zones (right)"}
p1 <- rowMassPlot(data)
p2 <- rowMassPlot(data_ling)

p1 + plot_spacer() + p2 + plot_layout(widths = c(1, 0.1, 1))
```

### Subcorpora dimension

We consider four different sizes for subcorpora dimension (in view of normalization by column): number of documents (`nDoc`), total number of word-tokens in documents (`dimCorpus`), sum of keyword frequencies (column sum) in the TDM (`Csum`), maximum keyword frequency (maximum column frequency) in the TDM (`Mcf`). The first two are in the corpus metadata, the latter two can be computed from the TDM. The four measures are in general strongly correlated (except for `Mcf`). They can be visualized graphically as bar plots or line plots.

```{r subcorpora-size}
ind_d <- data$year_cols
Csum <- tdm[, ind_d] %>% apply(2, sum, na.rm = TRUE)
Mcf <- tdm[, ind_d] %>% apply(2, max, na.rm = TRUE)

d_size <- corpus_info[, c(1, 3, 2)] %>% cbind(Csum, Mcf)

datatable(
  d_size,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Subcorpora dimensions by year"
)

summary(d_size[, -1])
apply(d_size[, -c(1, 5)], 2, sum)
```

Identify proper rescaling before plotting:

```{r rescaling}
sc <- c(1, 10, 10, 1)
apply(d_size[, -1], 2, max) / sc
ymax <- max(apply(d_size[, -1], 2, max) / sc)
ymax
```

They can be visualized graphically as bar plots or line plots:

```{r colmass-plot, fig.width=8, fig.height=4, fig.cap="Subcorpora dimension: for each year, number of texts (dot-line), total number of word-tokens(/10) in texts, column sum(/10) in the TDM, maximum column frequency in the TDM"}
colMassPlot(
  data, 
  sc = c(1, 10, 10, 1), 
  r = 1, 
  textty = "text", 
  themety = "light", 
  size_b = 2.5, 
  x_lab = "year"
)
```

### Temporal curves

The overall frequency trajectories (curves) are shown for the original data and separately for each frequency class or zone (with a sample for each class).

The plots are based on the long-format TDM stored in `data$tdm_long`, which has `year` as a group variable and `frequency` as the associated numerical variable.

```{r tdm-long}
tdm_l <- data$tdm_long

datatable(
  tdm_l,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    scrollY = "400px",
    dom = 'frtip'
  ),
  caption = "Term-Document Matrix in long format"
)
```

```{r curves-plot, fig.width=8, fig.height=5, fig.cap="Keyword trajectories (original data)"}
curvePlot(data, r = 2, size_class = c(0.2, 0.3, 0.4, 0.5))
```

Select sample keywords for detailed visualization:

```{r sample-keywords}
set.seed(2222)
r <- 3
kw <- tdm_l %>% 
  group_by(zone) %>% 
  summarise(keyword = list(sample(unique(keyword), r))) %>%
  ungroup()
kw
```

```{r curves-ctu-plot, fig.width=8, fig.height=5, fig.cap="Keyword trajectories with example keywords highlighted"}
ctu_noun <- c("person", "object", "instruct", "incident")

curveCtuPlot(data, ctu_noun = ctu_noun, r = 2, size_class = NULL)
```

Top keywords by frequency class:

```{r top-keywords}
top_kw <- tdm %>% 
  group_by(zone) %>% 
  arrange(desc(tot_freq)) %>% 
  slice_head(n = 3) %>% 
  ungroup() %>% 
  select(zone, keyword, tot_freq)

datatable(
  top_kw,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    dom = 't'
  ),
  caption = "Top 3 keywords by frequency zone"
)
```

```{r facet-plot, fig.width=7.5, fig.height=6, fig.cap="Keyword trajectories by frequency class (original data). Group data scale on y-axis"}
facetPlot(
  data, 
  keyword_selection = list(type = "frequency", n = 3, kw.list = NULL),
  r = 4, 
  scales = "free", 
  leg = TRUE, 
  size_class = NULL
)
```

# Normalization

Different normalization strategies can be applied to the TDM:

## Normalization by column

Subcorpus total tokens (dimCorpus):

```{r norm-nc}
datnc <- normalization(data, normty = "nc", sc = 1000)
```

## Normalization by row and column

Chi-square-like normalization:

```{r norm-nchi}
datnchi <- normalization(data, normty = "nchi", sc = 1)
```

Max row frequency and subcorpus total tokens:

```{r norm-nM}
datnM <- normalization(data, normty = "nM", sc = 1000)
```

MinMax row frequency and subcorpus total tokens:

```{r norm-nmM}
datnmM <- normalization(data, normty = "nmM", sc = 1000)
```

## Non-linear normalization

Non-linear transformation: Difference from the minimum weights more when the distribution is positively asymmetric (p>0.5), vice versa, difference from the maximum weights more when the distribution is negatively asymmetric (p<0.5).

```{r norm-nnl}
datnnl <- normalization(data, normty = "nnl", sc = 1000, nnlty = "M")
```

Check normalization results:

```{r norm-results}
names(datnc)
datnc$norm
datnc$normty

datatable(
  head(datnc$tdm, 20),
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Normalized TDM (first 20 rows)"
)

datatable(
  head(datnc$tdm_long, 50),
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Normalized TDM in long format (first 50 rows)"
)
```

Asymmetry analysis for non-linear normalization:

```{r asymmetry}
names(datnnl)
temp <- datnnl$tdm[, c(1, 4)] %>% cbind(datnnl$p_asy)

datatable(
  head(temp, 20),
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'frtip'
  ),
  caption = "Asymmetry coefficients (first 20 rows)"
)
```

<!-- # Session Information -->

<!-- ```{r session-info} -->
<!-- sessionInfo() -->
<!-- ``` -->

